News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-08-06 rupnik2015.tex
complete pass

\begin{abstract}
Today we follow news which is distributed globally. Significant events are reported by different sources and in different languages. In this work, we address the problem of tracking and events in a large multilingual stream. We consider a particular aspect of this problem, namely how to link collections of articles in different languages which refer to the same event. Given a multi-lingual stream and clusters of articles from each language, we first propose a method for cross-lingual document similarity based on Wikipedia, which enables us to compute the similarity of any two articles regardless of language. The proposed method can scale to 100 languages and can match articles from languages with little or no direct overlap in the training data. Using this similarity, we then propose an approach to link clusters of articles across languages which represent the same event. We provide an extensive evaluation of the system as a whole, as well as an evaluation of the quality and robustness of the similarity measure and the linking algorithm.
\begin{comment}
Main points:
\begin{itemize}
\item Problem of tracking events in a large multilingual stream - we reduce it to event linking
\item Two contributions. We propose a method that enables cross-lingaul document similarity an approach to linking multilingual data and an approach that uses is to link events.
\item Evaluation of cross-lingual approaches on Wikipedia and the event linking task.
\end{itemize}
\end{comment}
\end{abstract}
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item Cross lingual document similarity (model)
%\item  Event Detection
%\item  Combing the approaches
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would liek to understand events ans sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%%\end{itemize}
%
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item System overview
%\item Cross lingual document similarity (algorithms)
%\item Event linking
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would like to understand global news as sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%\end{itemize}
%

@@

Content on the internet is becoming increasingly multi-lingual. A prime example is Wikipedia - in 2001 the majority of pages were written in English, while in 2015, the percentage of English articles has dropped to 14\%. Machine translation remains relatively rudimentary - allowing people to understand simple phrases on web pages, but remain inadequate for more advanced understanding of text. At the same time, online news has begun to dominate reporting of current events. In this paper we consider the intersection of these developments: how to track events which are reported about in multiple languages.

@@

We base our techniques of cross-lingual event linking on an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles that are collected from numerous news outlets all over the world. The important components in the pipeline of the Event Registry are shown in Figure~\ref{fig:erpipeline} and will now be briefly described.

@@

The collection of the news articles is performed using the Newsfeed service~\cite{Trampus2012}. The service monitors RSS feeds of around 100.000 mainstream news outlets available globally. Whenever a new article is detected in the RSS feed, the service downloads all available information about the article and sends the article through the pipeline. Newsfeed downloads daily on average around 200.000 news articles in various languages, where English, Spanish and German are the most common.

@@

As a next step, an online clustering algorithm is applied on the articles in order to identify groups of articles that are discussing the same event. For each new article, the clustering algorithm determines if the article should be assigned to some existing cluster or into a new cluster. The underlying assumption is that articles that are describing the same event are similar enough and will therefore be put into the same cluster. For clustering, each new article is first tokenized, stop words are removed and words are stemmed.
The remaining tokens are represented in a vector-space model and normalized using TF-IDF (see Section~\ref{sec:tfidf} for the definition). The cosine similarity is used to find the most similar existing cluster, by comparing the document's vector to the centroid vector of each cluster. A user-defined threshold is used to determine if the article is not similar enough to any existing clusters (0.4 was used in our experiments).  If the highest similarity is above the threshold, the article is assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article. Whenever an article is assigned to a cluster, the cluster's centroid vector is also updated. Since articles about an event are commonly written only for a short period of time, we remove clusters once the oldest article in the cluster becomes more than 4 days old. This housekeeping mechanism prevents the clustering from becoming slow and also ensures that articles are not assigned to obsolete clusters. Details of the clustering approach are described in~\cite{brank2014}.

Once the number of articles in a cluster reaches a threshold (which is a language dependent parameter), we assume that the articles in the cluster are describing an event. At that point, a new event with a unique ID is created in Event Registry, and the cluster with the articles is assigned to it. By analyzing the articles we extract the main information about the event, such as the event location, date, most relevant entities and keywords, etc.

Since articles in a cluster are in a single language, we also want to identify any other existing clusters that are reporting about the same event in another language and join these clusters into the same event. This task is performed using a classification approach that is the second major contribution of this paper. It is described in details in Section~\ref{sec:linking}.

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day - implying that similarity computation should be highly scalable.  %The system should support as many languages as possible.
 We focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This  simplifies implementation and system maintenance.
To summarize, the following properties are desirable:
\begin{itemize}
\item \textbf{Training} - The training (building cross-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpus (as opposed to parallel translation based corpora), with missing data. Supporting a new language should be straightforward.
\item \textbf{Operation efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day.
\item \textbf{Operation cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is straightforward to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

@@

\textbf{Classification problem related}. In approaches such as \cite{cotrain} and \shortcite{svm2k} the authors find linear embeddings of the data that support good classification results, based on a given classification problem training data. We do not consider this family of approaches, since for general language pairs classifier training sets may not be available.

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CLLSI\shortcite{cl_lsi}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CLLSI, which will be presented in more detail.

@@

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are almost no services that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but they offer no linking of clusters across languages. The only service that we found, which provides cross-lingual cluster linking, is the European Media Monitor (EMM)~\cite{pouliquen2008story,Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\shortcite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists. If the similarity is above the threshold, the clusters are linked. Compared to EMM, our approach uses document similarities to obtain a small set of potentially equivalent clusters. Additionally, we do not decide if two clusters are equivalent based on a hand-set threshold on a similarity value -- instead we use a classification model that uses a larger set of features related to the tested pair of clusters.

A system that is significantly different, but worth mentioning is the GDELT project~\cite{Leetaru2013Gdelt}. In GDELT, events are also extracted from articles, but in their case, an event is specified in a form of a triple containing two actors and a relation. The project contains an extensive vocabulary of possible relations, mostly related to political events. In order to identify events, GDELT collects articles in more than 65 languages and uses machine translation to translate them to English. All information extraction is then done on the translated article.

\section{Cross-lingual Document Similarity}
\label{sec:crosslingual}
Document similarity is an important component in techniques from text mining and natural language processing. Many techniques use the similarity as a black box, i.e., a kernel in Support Vector Machines. Comparison of documents (or other types of text snippets) in a monolingual setting is a well studied problem in the field of information retrieval \cite{Salton88term-weightingapproaches}. We first recount some system requirements, then review  related approaches to cross-lingual similarity models and then formally introduce the problem as well as our approach.

\subsection{Problem definition}\label{sec:tfidf}
We will first describe how documents are represented as vectors and how to compare douments in a mono-lingual setting. We will then define a way to measure cross-lingual similarity that is natural for the models we consider.

\textbf{Document representation.}
The standard vector space model \cite{Salton88term-weightingapproaches} represents documents as vectors, where each term corresponds to word or phrase in a fixed vocabulary. Formally, document $d$ is represented by a vector $x \in \RR^n$, where $n$ corresponds to the size of the vocabulary, and vector elements $x_k$ correspond to the number of times term $k$ occurred in the document, also called \emph{term frequency} or $TF_k(d)$.

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$.  Therefore we can definte a document's $TFIDF$ as
$$ x_{ij}  := \frac{\mbox{term frequency in document } i}{\mbox{inverse document frequency of term } j}.$$
The $TFIDF$ weighted vector space model document representation corresponds to a map $\phi : \text{text} \rightarrow \RR^n$ defined by:
$$\phi(d)_k = {TF}_k(d) \log\left( \frac{N}{{DF}_k}\right).$$

\textbf {Mono-lingual similarity.}
A common way of computing similarity between documents is \emph{cosine similarity},
$$sim(d_1, d_2) = \frac{\langle \phi(d_1), \phi(d_2)\rangle}{\|\phi(d_1)\| \|\phi(d_2)\|},$$
where $\langle \cdot,\cdot \rangle$ and $\|\cdot\|$ are standard inner product and Euclidean norm. Cosine similarity, and other related approaches, assumes that the similarity is reflected in the overlap of words. This metric is only meaningful when the documents $d_1$ and $d_2$ are defined on the same space (e.g. have a comparable vocabulary or are written in the same language).

\textbf {Cross-lingual similarity.}
Processing a multilingual dataset results in several vector spaces with varying dimensionality, one for each language. The dimensionality of the vector space corresponding to the $i$-th language is denoted by $n_i$ and and the vector space model mapping is denoted by $\phi_i : \text{text} \rightarrow \RR^{n_i}$.
The similarity between documents in language $i$ and language $j$ is defined as a bilinear operator represented as a matrix $S_{i,j} \in \RR^{n_i \times n_j}$:
$$sim_{i,j}(d_1, d_2) = \frac{ \langle \phi_i (d_1), S_{i,j} \phi_j (d_2) \rangle }{\|\phi_i(d_1)\| \|\phi_j(d_2)\|},$$
where $d_1$ and $d_2$ are documents written in the $i$-th and $j$-th language respectively. If the the maximal singular value of $S_{i,j}$ is bounded by $1$, then the similarity scores will lie on the interval $[-1, 1]$. In Section \ref{sec:models} we will describe some approaches to computing $S_{i,j}$ given training data.

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we describe several approaches to the problem of computing multilingual simliarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new, more generally applicable method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.

CCA can be used to find
correlated patterns for a pair of languages, whereas the extended method optimizes a
sum of squared correlations between several language pairs. The method takes advantage of the hub language assumption in the training dataset (the English Wikipedia corresponds to a hub) as well as a preprocessing step based on LSI that acts as regularization and improves scalability.

@@

The cross-lingual similarity models presented in this paper are based on comparable corpora. That is, a corpus of documents in multiple languages, with alignment between documents that are of the same topic, or even a rough translation of each other. Wikipedia is an example of a comparable corpus, where a specific entry can be described in multiple languages (e.g. "Berlin" is currently described in 222 languages). News articles represent another example, where the same event can be described by newspapers in several languages.

More formally, \emph{multilingual document} $d = (u_1,\ldots u_m)$ is a tuple of $m$ documents on the same topic (comparable), where $u_i$ is the document written in language $i$. Note that individual document $u_i$ can be an empty document (missing resource) and each $d$ must contain at least two nonempty documents. A comparable corpus $D = {d_1, \ldots, d_N}$ is a collection of multilingual documents. By using the vector space model we can represent $D$ as a set of $m$ matrices $X_1,\ldots,X_m$, where $X_i \in \RR^{n_i \times N}$ is the matrix corresponding to the language $i$ and $n_i$ is the vocabulary size of language $i$. Furthermore, let $X_i^{\ell}$ denote the $\ell$-th column of matrix $X_i$ and the matrices respect the document alignment - the vector $X_i^\ell$ corresponds to the TFIDF vector of the $i$-th component of multilingual document $d_\ell$. We use $N$ to denote the total row dimension of $X$, i.e. $N:= \sum_{i=1}^m n_i$.

@@

The $k$-means algorithm is perhaps the most well-known and widely-used clustering algorithm. We now present its application
to computing cross-lingual similarities. The idea is based on concatenating the corpus matrices, running standard $k$-means clustering to obtain the matrix of centroids, "reversing" the concatenation step to obtain a set of aligned bases, which are finally used to compute cross-lingual similarities.

 In order to apply the algorithm, we first merge all the term-document matrices into a single matrix $X$ by stacking the individual term-document matrices:
$$X := \begin{bmatrix}X_1^T ,X_2^T, \cdots, X_m^T \end{bmatrix}^T,$$
such that the columns respect the alignment of the documents (here MATLAB notation for concatenating matrices is used). Therefore, each document  is represented by a long vector indexed by the terms in all languages.

We then run the $k$-means algorithm \cite{kmeans} and obtain a centroid matrix $C \in \RR^{N \times k}$, where the $k$ columns represent centroid vectors. The centroid matrix can be split vertically into $m$ blocks: $$C = [C_1^T \cdots C_m^T]^T,$$ according to the number of dimensions of each language, i.e. $C_i \in \RR^{n_i \times k}$.

The matrices $C_i$ are obtained by analyzing the multilingual corpus matrix $X$ (obtained from Wikipedia for example) and we now show how to use them to compute cross-lingual document similarities on new documents (news documents, for example).

Each matrix $C_i$ represents a vector space basis and can be used to map points in $\RR^{n_i}$ into a $k$-dimensional space, where the coordinates of a vector $x \in \RR^{n_i}$ are expressed as: $$(C_i^T C_i)^{-1} C_i^T x_i.$$

@@

The next method is Cross-Lingual Latent Semantic Indexing (CL-LSI)\cite{cl_lsi} which is a variant of LSI~\shortcite{lsi} for more than one language. The method is based on computing a truncated singular value decomposition of $X \approx U S V^T$. Since the matrix can be large we can use an iterative method like the Lanczos~\cite{golub} algorithm with reorthogonalization to find the left singular vectors (columns of $U$) corresponding to the largest singular values. It turns out that the Lanczos method converges slowly as the gap between the leading singular values is small. Moreover, the Lanczos method is hard to parallelize. Instead we use a randomized version of the singular value decomposition (SVD) described in \shortcite{tropp} that can be viewed as a block Lanczos method. That enables us to use parallelization and speeds up the computation considerably.

@@

\subsection{Canonical Correlation Analysis}\label{sec:CCA}
 The final approach is a statistical technique to analyze data from two sources - after which we will describe our main contribution to computing cross-lingual similarities.

 Canonical Correlation Analysis (CCA)~\cite{Hotelling} is a dimensionality reduction technique similar to Principal Component Analysis (PCA)~\cite{Pearson1901On}, with an additional assumption that the data consists of feature vectors that arose from two sources (two views) that share some information. Examples include: bilingual document collection ~\shortcite{mrpqr} and collection of images and captions~\shortcite{Hardoon_usingimage}. Instead of looking for linear combinations of features that maximize the variance (PCA) we look for a linear combination of feature vectors from the first view and a linear combination for the second view, that are maximally correlated.

>>>>

\begin{abstract}

In today's world, we follow news which is distributed globally. Significant events are reported by different sources and in different languages. In this work, we address the problem of tracking and events in a large multilingual stream. Within a recently developed system (EventRegistry  \shortcite{Leban2014W,Leban2014I}), we examine two aspects of this problem: how to compare articles in different languages and how to link collections of articles in different languages which refer to the same event.  Taking a multi-lingual stream and clusters of articles from each language, we compare different cross-lingual document similarity measures based on Wikipedia. This allows us to compute the similarity of any two articles regardless of language. Building on previous work, we show there are methods which scale well and can compute a meaningful similarity between articles from languages with little or no direct overlap in the training data.  
Using this capability, we then propose an approach to link clusters of articles across languages which represent the same event. We provide an extensive evaluation of the system as a whole, as well as an evaluation of the quality and robustness of the similarity measure and the linking algorithm.

%Given a multi-lingual stream and clusters of articles from each language, we first propose a method for cross-lingual document similarity based on Wikipedia, which enables us to compute the similarity of any two articles regardless of language. The proposed method can scale to 100 languages and can match articles from languages with little or no direct overlap in the training data. Using this similarity, we then propose an approach to link clusters of articles across languages which represent the same event. We provide an extensive evaluation of the system as a whole, as well as an evaluation of the quality and robustness of the similarity measure and the linking algorithm.
\begin{comment}
Main points:
\begin{itemize}
\item Problem of tracking events in a large multilingual stream - we reduce it to event linking
\item Two contributions. We propose a method that enables cross-lingaul document similarity an approach to linking multilingual data and an approach that uses is to link events.
\item Evaluation of cross-lingual approaches on Wikipedia and the event linking task.
\end{itemize}
\end{comment}
\end{abstract}
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item Cross lingual document similarity (model)
%\item  Event Detection
%\item  Combing the approaches
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would liek to understand events ans sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%%\end{itemize}
%
%\section{Overview}
%Structure
%\begin{itemize}
%\item Introduction
%\item System overview
%\item Cross lingual document similarity (algorithms)
%\item Event linking
%\item Experiments
%\item Conclusions
%\end{itemize}
%The story:
%\begin{itemize}
%\item Description of the problem - we would like to understand global news as sequences of events, but our observables (e.g. news stories) occur in a variety of languages
%\item We present an approach to tackle this problem using a combination of techniques....
%\item Ultimately we present a tool showing the feasibility of our approach
%\end{itemize}
%

@@

Content on the internet is becoming increasingly multi-lingual. A prime example is Wikipedia - in 2001, the majority of pages were written in English, while in 2015, the percentage of English articles has dropped to 14\%. At the same time, online news has begun to dominate reporting of current events. However, machine translation remains relatively rudimentary - allowing people to understand simple phrases on web pages, but remain inadequate for more advanced understanding of text. In this paper we consider the intersection of these developments: how to track events which are reported about in multiple languages.

@@

We base our techniques of cross-lingual event linking on an online system for detection of world events, called Event Registry \shortcite{Leban2014W,Leban2014I}. Event Registry is a repository of events, where events are automatically identified by analyzing news articles that are collected from numerous news outlets all over the world. The important components in the pipeline of the Event Registry are shown in Figure~\ref{fig:erpipeline} and are briefly described below.

@@

The collection of the news articles is performed using the Newsfeed service~\cite{Trampus2012}. The service monitors RSS feeds of around 100,000 mainstream news outlets available globally. Whenever a new article is detected in the RSS feed, the service downloads all available information about the article and sends the article through the pipeline. Newsfeed downloads daily on average around 200,000 news articles in various languages, where English, Spanish and German are the most common.

@@

As a next step, an online clustering algorithm is applied to the articles in order to identify groups of articles that are discussing the same event. For each new article, the clustering algorithm determines if the article should be assigned to some existing cluster or into a new cluster. The underlying assumption is that articles that are describing the same event are similar enough and will therefore be put into the same cluster. For clustering, each new article is first tokenized, stop words removed and the remaining words are stemmed.
The remaining tokens are represented in a vector-space model and normalized using TF-IDF (see Section~\ref{sec:tfidf} for the definition). Cosine similarity is used to find the most similar existing cluster, by comparing the document's vector to the centroid vector of each cluster. A user-defined threshold is used to determine if the article is not similar enough to any existing clusters (0.4 was used in our experiments).  If the highest similarity is above the threshold, the article is assigned to the corresponding cluster, otherwise a new cluster is created, initially containing only the single article. Whenever an article is assigned to a cluster, the cluster's centroid vector is also updated. Since articles about an event are commonly written only for a short period of time, we remove clusters once the oldest article in the cluster becomes more than 4 days old. This housekeeping mechanism prevents the clustering from becoming slow and also ensures that articles are not assigned to obsolete clusters. Details of the clustering approach are described in~\cite{brank2014}.

Once the number of articles in a cluster reaches a threshold (which is a language dependent parameter), we assume that the articles in the cluster are describing an event. At that point, a new event with a unique ID is created in Event Registry, and the cluster with the articles is assigned to it. By analyzing the articles, we extract the main information about the event, such as the event location, date, most relevant entities and keywords, etc.

Since articles in a cluster are in a single language, we also want to identify any other existing clusters that report about the same event in other languages and join these clusters into the same event. This task is performed using a classification approach that is the second major contribution of this paper. It is described in detail in Section~\ref{sec:linking}.

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day. Both step rely heavily on simliarity computation, which implies that this must be highly scalable. 
%
%while both steps must be scalable, the underlying similarity computation is used 
%- implying that similarity computation should be highly scalable.  %The system should support as many languages as possible.
Therefore, we focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This  simplifies implementation and system maintenance.
To summarize, the following properties are desirable:
\begin{itemize}
\item \textbf{Training} - The training (building cross-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpus (as opposed to parallel translation based corpora), with missing data. Supporting a new language should be straightforward.
\item \textbf{Operation efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day.
\item \textbf{Operation cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is straightforward to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

@@

\textbf{Classification problem related}. In approaches such as \cite{cotrain} and \shortcite{svm2k} the authors find linear embeddings of the data that support good classification results, based on a given classification problem training data. We do not consider this family of approaches, since for general language pairs classifier training sets may not be available.\todo{add more here}

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CLLSI\shortcite{cl_lsi}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CLLSI.

\todo{add the related work which was missing}

@@

Although there are a number of services that aggregate news by identifying clusters of similar articles, there are almost no services that provide linking of clusters over different languages. Google News as well as Yahoo! News are able to identify clusters of articles about same event, but offer no linking of clusters across languages. The only service that we found, which provides cross-lingual cluster linking, is the European Media Monitor (EMM)~\cite{pouliquen2008story,Steinberger2008}. EMM clusters articles in 60 languages and then tries to determine which clusters of articles in different languages describe the same event. To achieve cluster linking, EMM uses three different language independent vector representations for each cluster. First vector contains the weighted list of references to countries mentioned in the articles, while the second vector contains the weighted list of mentioned people and organizations. The last vector contains the weighted list of Eurovoc subject domain descriptors. These descriptors are topics, such as \emph{air transport}, \emph{EC agreement}, \emph{competition} and \emph{pollution control} into which articles are automatically categorized~\shortcite{Pouliquen2006}. Similarity between clusters is then computed using a linear combination of the cosine similarities computed on the lists. If the similarity is above the threshold, the clusters are linked. Compared to EMM, our approach uses document similarities to obtain a small set of potentially equivalent clusters. Additionally, we do not decide if two clusters are equivalent based on a hand-set threshold on a similarity value -- instead we use a classification model that uses a larger set of features related to the tested pair of clusters.

A system that is significantly different, but worth mentioning is the GDELT project~\cite{Leetaru2013Gdelt}. In GDELT, events are also extracted from articles, but in their case, an event is specified in a form of a triple containing two actors and a relation. The project contains an extensive vocabulary of possible relations, mostly related to political events. In order to identify events, GDELT collects articles in more than 65 languages and uses machine translation to translate them to English. All information extraction is then done on the translated article.
\todo{add the missing related work}

\section{Cross-lingual Document Similarity}
\label{sec:crosslingual}
Document similarity is an important component in techniques from text mining and natural language processing. Many techniques use the similarity as a black box, e.g., a kernel in Support Vector Machines. Comparison of documents (or other types of text snippets) in a monolingual setting is a well studied problem in the field of information retrieval \cite{Salton88term-weightingapproaches}. We first formally introduce the problem followed by our approach.

\subsection{Problem definition}\label{sec:tfidf}
We will first describe how documents are represented as vectors and how to compare douments in a mono-lingual setting. We then define a way to measure cross-lingual similarity that is natural for the models we consider.

\noindent\textbf{Document representation.}
The standard vector space model \cite{Salton88term-weightingapproaches} represents documents as vectors, where each term corresponds to word or phrase in a fixed vocabulary. Formally, document $d$ is represented by a vector $x \in \RR^n$, where $n$ corresponds to the size of the vocabulary, and vector elements $x_k$ correspond to the number of times term $k$ occurred in the document, also called \emph{term frequency} or $TF_k(d)$.

We also used a term re-weighting scheme that adjusts for the fact that some words occur more frequently in general. A term weight should correspond to the importance of the term for the given corpus. The common weighting scheme is called \emph{Term Frequency Inverse Document Frequency} ($TFIDF$) weighting. An \emph{Inverse Document Frequency} ($IDF$) weight for the dictionary term $k$ is defined as $\log\left( \frac{N}{DF_k} \right)$, where $DF_k$ is the number of documents in the corpus which contain term $k$. \todo{our corpus is dynamic how do we deal with this?} Therefore we can definte a document's $TFIDF$ as
$$ x_{ij}  := \frac{\mbox{term frequency in document } i}{\mbox{inverse document frequency of term } j}.$$
The $TFIDF$ weighted vector space model document representation corresponds to a map $\phi : \text{text} \rightarrow \RR^n$ defined by:
$$\phi(d)_k = {TF}_k(d) \log\left( \frac{N}{{DF}_k}\right).$$

\noindent\textbf {Mono-lingual similarity.}
A common way of computing similarity between documents is \emph{cosine similarity},
$$sim(d_1, d_2) = \frac{\langle \phi(d_1), \phi(d_2)\rangle}{\|\phi(d_1)\| \|\phi(d_2)\|},$$
where $\langle \cdot,\cdot \rangle$ and $\|\cdot\|$ are standard inner product and Euclidean norm. Cosine similarity, and other related approaches, assumes that the similarity is reflected in the overlap of words. While always defined, this metric is only meaningful when the documents $d_1$ and $d_2$ are defined on the same space (e.g. have a comparable vocabulary or are written in the same language, since otherwise their supports are disjoint).

\noindent\textbf {Cross-lingual similarity.}
Processing a multilingual dataset results in several vector spaces with varying dimensionality, one for each language. The dimensionality of the vector space corresponding to the $i$-th language is denoted by $n_i$ and and the vector space model mapping is denoted by $\phi_i : \text{text} \rightarrow \RR^{n_i}$.
The similarity between documents in language $i$ and language $j$ is defined as a bilinear operator represented as a matrix $S_{i,j} \in \RR^{n_i \times n_j}$:
$$sim_{i,j}(d_1, d_2) = \frac{ \langle \phi_i (d_1), S_{i,j} \phi_j (d_2) \rangle }{\|\phi_i(d_1)\| \|\phi_j(d_2)\|},$$
where $d_1$ and $d_2$ are documents written in the $i$-th and $j$-th language respectively. If the the maximal singular value of $S_{i,j}$ is bounded by $1$, then the similarity scores will lie on the interval $[-1, 1]$. In Section \ref{sec:models} we will describe some approaches to computing $S_{i,j}$ given training data. \todo{add some more details/intuition.}

\subsection{Cross-Lingual Models}\label{sec:models}
In this section, we describe several approaches to the problem of computing multilingual simliarities introduced in \ref{sec:tfidf}. We present four approaches:
a simple approach based on $k$-means clustering in \ref{sec:kmeans}, a standard approach based on singular value decomposition in \ref{sec:LSI}, a related
approach called Canonical Correlation Analysis (CCA) in \ref{sec:CCA} and finally a new, more generally applicable method, which is an extension of CCA to more than two languages in \ref{sec:hublang}.
%
CCA can be used to find
correlated patterns for a pair of languages, whereas the extended method optimizes a
sum of squared correlations between several language pairs. The method takes advantage of the hub language assumption in the training dataset (the English Wikipedia corresponds to a hub) as well as a preprocessing step based on LSI that acts as regularization and improves scalability. \todo{we need to set this up better}

@@

The cross-lingual similarity models presented in this paper are based on comparable corpora. That is, a corpus of documents in multiple languages, with alignment between documents that are of the same topic, or even a rough translation of each other. Wikipedia is an example of a comparable corpus, where a specific entry can be described in multiple languages (e.g. ``Berlin" is currently described in 222 languages). News articles represent another example, where the same event can be described by newspapers in several languages.

More formally, \emph{multilingual document} $d = (u_1,\ldots u_m)$ is a tuple of $m$ documents on the same topic (comparable), where $u_i$ is the document written in language $i$. Note that individual document $u_i$ can be an empty document (missing resource) and each $d$ must contain at least two nonempty documents. A comparable corpus $D = {d_1, \ldots, d_N}$ is a collection of multilingual documents. By using the vector space model we can represent $D$ as a set of $m$ matrices $X_1,\ldots,X_m$, where $X_i \in \RR^{n_i \times N}$ is the matrix corresponding to the language $i$ and $n_i$ is the vocabulary size of language $i$. Furthermore, let $X_i^{\ell}$ denote the $\ell$-th column of matrix $X_i$ and the matrices respect the document alignment - the vector $X_i^\ell$ corresponds to the TFIDF vector of the $i$-th component of multilingual document $d_\ell$. We use $N$ to denote the total row dimension of $X$, i.e. $N:= \sum_{i=1}^m n_i$.\todo{picture - this part isnt that clear}

@@

The $k$-means algorithm is perhaps the most well-known and widely-used clustering algorithm. Here, we present its application
to computing cross-lingual similarities. The idea is based on concatenating the corpus matrices, running standard $k$-means clustering to obtain the matrix of centroids, ``reversing" the concatenation step to obtain a set of aligned bases, which are finally used to compute cross-lingual similarities.

 In order to apply the algorithm, we first merge all the term-document matrices into a single matrix $X$ by stacking the individual term-document matrices:
$$X := \begin{bmatrix}X_1^T ,X_2^T, \cdots, X_m^T \end{bmatrix}^T,$$
such that the columns respect the alignment of the documents (here MATLAB notation for concatenating matrices is used). Therefore, each document  is represented by a long vector indexed by the terms in all languages. \todo{picture here}

We then run the $k$-means algorithm \cite{kmeans} and obtain a centroid matrix $C \in \RR^{N \times k}$, where the $k$ columns represent centroid vectors. The centroid matrix can be split vertically into $m$ blocks: $$C = [C_1^T \cdots C_m^T]^T,$$ according to the number of dimensions of each language, i.e. $C_i \in \RR^{n_i \times k}$.
%
The matrices $C_i$ are computed using a multilingual corpus matrix $X$ (obtained from Wikipedia for example).

To compute  cross-lingual document similarities on new documents, note that each matrix $C_i$ represents a vector space basis and can be used to map points in $\RR^{n_i}$ into a $k$-dimensional space, where the coordinates of a vector $x \in \RR^{n_i}$ are expressed as: $$(C_i^T C_i)^{-1} C_i^T x_i.$$

@@

The second approach we consider is Cross-Lingual Latent Semantic Indexing (CL-LSI)\cite{cl_lsi} which is a variant of LSI~\shortcite{lsi} for more than one language. The method is based on computing a truncated singular value decomposition of $X \approx U S V^T$. Since the matrix can be large we can use an iterative method like the Lanczos~\cite{golub} algorithm with reorthogonalization to find the left singular vectors (columns of $U$) corresponding to the largest singular values. It turns out that the Lanczos method converges slowly as the gap between the leading singular values is small. Moreover, the Lanczos method is hard to parallelize. Instead, we use a randomized version of the singular value decomposition (SVD) described in \shortcite{tropp} that can be viewed as a block Lanczos method. That enables us to use parallelization and speeds up the computation considerably.

@@

\subsection{Canonical Correlation Analysis}\label{sec:CCA}
 The final approach is a statistical technique to analyze data from two sources - after which we will describe our extension to computing cross-lingual similarities.
%
 Canonical Correlation Analysis (CCA)~\cite{Hotelling} is a dimensionality reduction technique similar to Principal Component Analysis (PCA)~\cite{Pearson1901On}, with an additional assumption that the data consists of feature vectors that arose from two sources (two views) that share some information. Examples include: bilingual document collection ~\shortcite{mrpqr} and collection of images and captions~\shortcite{Hardoon_usingimage}. Instead of looking for linear combinations of features that maximize the variance (PCA) we look for a linear combination of feature vectors from the first view and a linear combination for the second view, that are maximally correlated.
